Global patterns of forest autotrophic carbon fluxes

### 2020-08-06 manuscript/manuscript-draft.tex
working on discusion
(still a mess, pushing now to avoid conflicted copies)

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \usepackage{setspace}\onehalfspacing \usepackage{lineno}\linenumbers

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Conservation Ecology Center; Smithsonian Conservation Biology
  Institute; National Zoological Park, Front Royal, VA, USA
\item
  \emph{Becky- current}
\item
  Center for Tropical Forest Science-Forest Global Earth Observatory;
  Smithsonian Tropical Research Institute; Panama, Republic of Panama
\item
  Joint Global Change Research Institute, Pacific Northwest National
  Laboratory, College Park Maryland 20740 USA
\end{enumerate}

@@

fax:1-540-635-6506

@@

\subsubsection{Abstract}\label{abstract}

{[}\textbf{very rough start:}{]} Carbon fixation, allocation, and
metabolism by trees set the basis for energy and material flows in
forest ecosystems and define their interactions with Earth's changing
climate. Here, we drew upon \# records from the Global Forest Carbon
Database (ForC), representing all major forest types and the nine most
significant forest autotrophic carbon flux (FACF) variables, to
comprehensively explore how C cycling in mature, undisturbed forests
varies with latitude and climate on a global scale. We show that, across
all FACF variables analyzed, C cycling decreases linearly with latitude
-- a finding that confirms multiple previous studies but contradicts the
idea that net primary productivity (\(NPP\)) of temperate forests rivals
that of tropical forests. The FACF variables generally increase in
proportion to one another, with few differences in allocation detectable
at this global scale, but differed in that latitude explained a lower
proportion of variation among subsidiary fluxes (in particular, woody
aboveground \(NPP\) and belowground \(NPP\), \(BNPP\)). Climate
explained a significant proportion (\#-\#\%) of variation in all C
fluxes analyzed (less for subsidiary fluxes), with temperature variables
in general and mean annual temperature (\(MAT\)) in particular being the
best predictors of FACF on this global scale. While other climate
variables (\emph{e.g.}, \textbf{XX}) displayed significant correlation
with FACF, none of them had significantly better explanatory power than
\(MAT\). The effects of temperature were modified by moisture
availability, with reduced FACF under hot and dry conditions and
sometimes under very high precipitaiton (especially for \(BNPP\)). FACF
declined with temperature seasonality, but growing season length doesn't
improve upon MAT as a predictor. Within the growing season, the
influence of climate on C cycling is smaller but still significant for a
number of carbon fluxes. These findings clarify the big picture of how
FACF varies with latitude and climate on a global scale. As we enter a
period of accelerating climatic change, understanding of the fundamental
climatic controls on FACF sets a foundation for understanding patterns
of change.

@@

\subsubsection{Introduction}\label{introduction}

Carbon cycling in forests worldwide provides the energetic basis for
sustaining the majority of Earth's terrestrial biodiversity and many
human populations (\textbf{REF}), while strongly influencing atmospheric
CO\textsubscript{2} and climate \citep{bonan_forests_2008}. Forests'
autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation,
and metabolism by trees and other primary producers--sets the energy
ultimately available to heterotrophic organisms (including microbes), in
turn influencing their abundance (\textbf{REFS}) and possibly diversity
\citep{waide_relationship_1999} (\textbf{REFS}). FACF influences all
organic matter stocks in forest ecosystems and is linked to cycling of
energy, water, and nutrients (\textbf{REFS}) \citep{piao_forest_2010}.
Critically, FACF also define forest interactions with Earth's changing
climate. Over 69 Gt of CO\textsubscript{2} cycle through Earth's forests
each year \citep{badgley_terrestrial_2019}, and in recent decades their
net C sequestration (\textasciitilde{}2.4 Gt C yr\textsuperscript{-1})
offset roughly 30\% of anthropogenic fossil fuel emissions
\citep{pan_large_2011}. As atmospheric carbon dioxide levels continue to
rise, driving climate change, forests will play a critical role in
shaping the future of Earth's climate
\citep{cavaleri_urgent_2015, rogelj_mitigation_2018}. However, our
ability to draw general macroscopic conclusions regarding global
variation in FACF with respect to climate has been limited in that these
analyses often mix forests that vary in stand age, disturbance history,
and/or management status; do not always sufficiently parse related
variables; and typically consider only one or a few variables at a time.

FACF decrease with latitude, but it remains unclear whether and how the
shape of this relationship varies across fluxes. Studies agree that FACF
are lowest in the boreal regions, and increase into the temperate
regions
\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
However, evidence is inconclusive on whether primary productivity
continues to increase into the tropics, or whether it plateaus in
temperate regions. Evidence for this is further complicated by the fact
that different studies use different measures of productivity to explore
these relationships. For example, modelling of global terrestrial
ecosystem gross primary productivity (\(GPP\)) through upscaling and
calibration of eddy flux measurements indicates that GPP peaks in
tropical forests
\citep{beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
In contrast, some studies suggest that the highest values of net primary
productivity (\(NPP\)) may be found in temperate forests
\citep{luyssaert_co_2007, huston_global_2009}, while others find \(NPP\)
highest in the tropics and decreasing with latitude
\citep{simova_enigma_2017}. Other studies have chosen to focus
exclusively on above-ground net primary productivity (\(ANPP\)), finding
evidence of a weak negative relationship between \(ANPP\) and latitude
\citep{huston_global_2009, gillman_latitude_2015}.

The latitudinal gradient in FACF is primarily driven by climate, which
is a significant driver of FACF across broad spatial scales
\citep{luyssaert_co_2007, cleveland_relationships_2011, hursh_sensitivity_2017}.
The majority of studies have focused on exploring the relationships of
FACF to mean annual temperature (MAT) and mean annual precipitation
(MAP), as the most commonly reported site-level climate variables. While
these fail to capture some important aspects of climate such as
seasonality, they do describe broad trends in temperature and water
availability, and therefore capture a substantial portion of
global-scale variation in climate. There is strong evidence that both
MAT and MAP show significant positive relationships with FACF
\citep{chu_does_2016}. However, as with latitude, the shape of those
relationships is not always clear, and, again, is complicated by the use
of different measures of FACF across studies. Various measures of
primary productivity \{\textbf{FACF?}\} saturate at high levels of MAP,
though the saturation points identified vary from 1500mm
\citep{luyssaert_co_2007} up to 2445mm MAP
\citep{schuur_productivity_2003}. Studies of the influence of MAT on
productivity \{\textbf{FACF?}\} are less conclusive. Luyssaert et al.
\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
while \(GPP\) increases linearly with MAT, NPP saturates at around
10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
\citeyearpar{larjavaara_temperature_2012}, find that increases in
\(GPP\) saturate at approximately 25\(^\circ\)C MAT, while Schuur
\citeyearpar{schuur_productivity_2003} shows that \(NPP\) increases
linearly with temperature. Taylor et al.
\citeyearpar{taylor_temperature_2017} showed a positive interaction
between MAT and MAP in shaping tropical forest productivity, such that
high rainfall had a negative effect on productivity in cooler climates,
compared to a positive effect in warmer climates. \{\textbf{It would be
good to add some more citations on soil respiraiton. I'm sure BBL can
help.}\}

FACF can be influenced by many other factors as well, which often act
across a range of scales, and may show interactive effects with each
other \citep{cleveland_relationships_2011}. On a local scale, stand age
\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity
\citep{liang_positive_2016}, management \citep{simova_enigma_2017},
nutrient availability \citep{aragao_above_2009}, and altitude
\citep{girardin_net_2010, malhi_variation_2017} all impact FACF. On a
global scale, we expect that FACF are most strongly influenced by broad
climatic gradients. There is evidence that FACFs also respond to
variables such as cloud cover \citep{taylor_temperature_2017}, solar
radiation \citep{fyllas_solar_2017}, and potential evapotranspiration
\citep{kerkhoff_plant_2005} in potentially significant ways.
Furthermore, MAT and MAP are very coarse measures of climate, and so
fail to capture much variation in climate on an intra-annual scale,
including the effects of factors such as growing season length, number
of frost-free days, temperature seasonality, and dry season length. Some
studies have suggested that the apparently strong relationship between
MAT and FACFs is actually a consequence of the correlation between MAT
and growing season length
\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
\citeyearpar{michaletz_convergence_2014} find that, within the growing
season, there is no significant relationship between primary
productivity and MAT, suggesting that the effect of temperature is due
to increased length of growing season, rather than an inherent influence
of temperature on FACF.

The recent development of the Global Forest Carbon database (ForC),
which synthesizes multiple variables and including records of stand
history
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018},
opens up the possibility for a standardized analysis of global scale
variation in multiple FACF and the principle climatic drivers of these
patterns. In order to approach this broad topic, we simplify the major
gaps in our knowledge to five broad questions and corresponding
hypotheses (Table 1). First, we ask how FACF vary with latitude. We then
test how these fluxes relate to MAT and MAP, and additionally how they
respond to other, less well-studied, climate variables. Finally, we
consider the relationship between FACF and seasonality, considering the
role of seasonality in explaining variation in carbon fluxes, and the
influence of climate on FACF standardized by growing season length. We
address the above questions for nine carbon fluxes contained in ForC,
allowing for an in-depth exploration of the effect of climate on FACF
globally.

@@

\caption{\label{tab:unnamed-chunk-4}Summary of research questions, corresponding hypotheses, and results. Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', parentheses indicate partial overall support/rejection of hypotheses across all FACF, and '-' indicates no significant relationship.}

@@

\begin{tabular}{llllllllllll}

@@

\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes (FACF)} & \multicolumn{1}{c}{ } \\

@@

Questions and hypotheses (with related references) & Overall & $GPP$ & $NPP$ & $ANPP$ & $ANPP_{woody.stem}$ & $ANPP_{foliage}$ & $BNPP$ & $BNPP_{fine.root}$ & $R_{auto}$ & $R_{auto-root}$ & Support\\

@@

\multicolumn{5}{l}{\textbf{Q1. How do FACF vary with latitude?}}\\
\hspace{1em}H1.1. FACF decrease linearly with latitude.\textsuperscript{1}$^{,}$\textsuperscript{2}$^{,}$\textsuperscript{3} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\

@@

\multicolumn{5}{l}{\textbf{Q2. How do FACF vary with MAT and MAP?}}\\
\hspace{1em}H2.1. FACF increase with MAT.\textsuperscript{1}$^{,}$\textsuperscript{4} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H2.2. FACF increase with precipitation.\textsuperscript{1}$^{,}$\textsuperscript{4} & (yes) & yes & yes & yes & - & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H2.3. Temperature and precipitation interactively shape FACF.\textsuperscript{5} & (yes) & yes & yes & - & yes & - & yes & yes & yes & - & Fig. 3\\

@@

\multicolumn{5}{l}{\textbf{Q3. How are FACF related to other climate variables?}}\\
\hspace{1em}H3.1. FACF increase with PET. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H3.2. FACF increase with vapour pressure deficit. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H3.3. FACF increase with solar radiation. & (yes) & yes & yes & yes & yes & yes & yes & yes & yes & - & Figs. S4, S5\\

@@

\multicolumn{5}{l}{\textbf{Q4. How does seasonality influence FACF?}}\\
\hspace{1em}H4.1. FACF decrease with temperature seasonality. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S6, S7\\
\hspace{1em}H4.2. FACF decrease with precipitation seasonality. & - & - & - & - & - & - & - & - & - & - & Figs. S6, S7\\
\hspace{1em}H4.3. FACF increase with growing season length.\textsuperscript{6}$^{,}$\textsuperscript{7}$^{,}$\textsuperscript{8} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S6, S7\\
\hspace{1em}H4.4. Growing season length is a better predictor of FACF than MAT.\textsuperscript{7}$^{,}$\textsuperscript{8} & (no) & no & no & no & no & no & no & - & no & no & Table S4\\

@@

\multicolumn{5}{l}{\textbf{Q5. When standardised by growing season length, how do FACF vary with climate?}}\\
\hspace{1em}H5.1. Growing season FACF increase with temperature.\textsuperscript{8} & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
\hspace{1em}H5.2. Growing season FACF increase with PET. & (yes) & yes & yes & - & yes & - & yes & yes & - & - & Figs. S8, S9\\
\hspace{1em}H5.3. Growing season FACF increase with precipitation. & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
\hspace{1em}H5.4. Growing season FACF increase with solar radiation. & (yes) & yes & yes & - & - & - & yes & yes & - & - & Figs. S8, S9\\

@@

\multicolumn{12}{l}{\textsuperscript{1} Luyssaert et al. (2007) \textsuperscript{2} Gillman et al. (2015) \textsuperscript{3} Simova and Storch (2017) \textsuperscript{4} Schuur (2003) \textsuperscript{5} Taylor et al. (2016) \textsuperscript{6} Malhi (2012) \textsuperscript{7} Michaletz et al. (2014) \textsuperscript{8} Chu et al. (2016)}\\

@@

\subsubsection{Materials and Methods}\label{materials-and-methods}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \usepackage{setspace}\onehalfspacing \usepackage{lineno}\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[]{natbib}
\bibliographystyle{apalike}

\author{}
\date{\vspace{-2.5em}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Conservation Ecology Center; Smithsonian Conservation Biology
  Institute; Front Royal, VA, USA
\item
  \emph{Becky- current}
\item
  Center for Tropical Forest Science-Forest Global Earth Observatory;
  Smithsonian Tropical Research Institute; Panama, Republic of Panama
\item
  Joint Global Change Research Institute, Pacific Northwest National
  Laboratory, College Park Maryland 20740 USA
\end{enumerate}

@@

\url{fax:1-540-635-6506}

@@

\#\#\#Abstract Carbon (C) fixation, allocation, and metabolism by trees
set the basis for energy and material flows in forest ecosystems and
define their interactions with Earth's changing climate. However, we
lack a cohesive synthesis on how forest carbon fluxes vary globally with
respect to climate and one another. Here, we draw upon 1319 records from
the Global Forest Carbon Database (ForC), representing all major forest
types and the nine most significant autotrophic carbon fluxes, to
comprehensively explore how C cycling in mature, undisturbed forests
varies with latitude and climate on a global scale. We show that, across
all flux variables analyzed, C cycling decreases continuously with
absolute latitude -- a finding that confirms multiple previous studies
but contradicts the idea that net primary productivity (\(NPP\)) of
temperate forests rivals that of tropical forests. C flux variables
generally displayed similar trends across latitude and multiple climate
variables, with no differences in allocation detected at this global
scale. Temperature variables in general, and mean annual temperature
(\(MAT\)) and temperature seasonality in particular, were the best
univariate predictors of C flux, explaining 19 - 71\% of variation in
the C fluxes analyzed. The effects of temperature were modified by
moisture availability, with C flux reduced under hot and dry conditions
and sometimes under very high precipitation. C fluxes increased with
growing season length, but this was never the best univariate predictor.
Within the growing season, the influence of climate on C cycling was
small but significant for a number of flux variables. These findings
clarify how forest C flux varies with latitude and climate on a global
scale. In a period of accelerating climatic change, this improved
understanding of the fundamental climatic controls on forest C cycling
sets a foundation for understanding patterns of change.

@@

\hypertarget{introduction}{%
\subsubsection{Introduction}\label{introduction}}

Carbon (C) cycling in Earth's forests provides the energetic basis for
sustaining the majority of Earth's terrestrial biodiversity and many
human populations
\citep{millennium_ecosystem_assessment_ecosystems_2005}, while strongly
influencing atmospheric carbon dioxide (CO\textsubscript{2}) and climate
\citep{bonan_forests_2008}. Forests' autotrophic carbon fluxes--that is,
carbon fixation, allocation, and metabolism by trees and other primary
producers--sets the energy ultimately available to heterotrophic
organisms (including microbes), in turn influencing their abundance
\citep{zak_plant_1994, niedzialkowska_species_2010} and possibly
diversity \citep{waide_relationship_1999, chu_direct_2018}. They are
linked to cycling of energy, water, and nutrients, and, critically,
influence all C stocks and define forest interactions with Earth's
changing climate. Each year, over 69 Gt of C cycle through Earth's
forests \citep{badgley_terrestrial_2019}-- a flux more than seven times
greater that of recent anthropogenic fossil fuel emissions \citep[9.5 Gt
C yr\textsuperscript{-1};][]{friedlingstein_global_2019}. As atmospheric
CO\textsubscript{2} continues to rise, driving climate change, forests
will play a critical role in shaping the future of Earth's climate
\citep{cavaleri_urgent_2015, rogelj_mitigation_2018}. However, our
understanding of the climate dependence of forest C cycling on a global
scale has been limited by analyses typically considering only one or a
few variables at a time, insufficient parsing of related variables, and
the mixing of data from forests that vary in stand age, disturbance
history, and/or management status, all of which affect C cycling
\citep{litton_carbon_2007, gillman_latitude_2015, simova_enigma_2017}.

Forest C fluxes decrease with latitude
\citep[e.g.,][]{luyssaert_co_2007, piao_forest_2010, gillman_latitude_2015, li_mapping_2019},
but it remains unclear whether and how the shape of this relationship
varies among fluxes. For instance, gross primary productivity (\(GPP\))
increases continuously with decreasing latitude and is indisputably
highest in tropical forests
\citep{luyssaert_co_2007, beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
In contrast, some studies have suggested that net primary productivity
(\(NPP\)), or its aboveground portion (\(ANPP\)), exhibits a less
distinct increase from temperate to tropical forests
\citep{luyssaert_co_2007}--or even a decrease \citep[but see
\citet{gillman_latitude_2015}]{huston_global_2009}. A shallower increase
in \(NPP\) than in \(GPP\) with decreasing latitude would align with the
suggestion that tropical forests tend to have low carbon use efficiency
\citep[\(CUE\)=
\(NPP\)/\(GPP\);][]{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
However, accurately comparing relationships of multiple forest C fluxes
across latitude requires a large database with standardization for
methodology, stand history, and management, which to date has not been
possible.

The latitudinal gradient in forest C flux rates, along with altitudinal
gradients \citep{girardin_net_2010, malhi_variation_2017}, is driven
primarily by climate, which is a significant driver of C fluxes across
broad spatial scales
\citep{luyssaert_co_2007, cleveland_relationships_2011, hursh_sensitivity_2017}.
However, there is little consensus as to the shapes of these
relationships or the best predictor variables. The majority of studies
have focused on exploring the relationships of C fluxes to mean annual
temperature (MAT) and mean annual precipitation (MAP), as the most
commonly reported site-level climate variables. C fluxes increase
strongly with \(MAT\) on the global scale, but whether they saturate or
potentially decrease at higher temperatures remains disputed. Some
studies have detected no deceleration or decline in \(GPP\)
\citep{luyssaert_co_2007} or \(NPP\) \citep{schuur_productivity_2003}
with increasing MAT. Some have suggested that relationship between C
flux and temperature is exponential in nature \citep{allen_linking_2005}
(\textbf{REFS}). In contrast, others have found evidence of saturation
or decline of C flux in the warmest climates; \citet{luyssaert_co_2007}
found \(NPP\) saturating at around 10\(^\circ\)C MAT;
\citet{larjavaara_temperature_2012} found that increases in \(GPP\)
saturate at approximately 25\(^\circ\)C MAT, and
\citet{sullivan_sensitivity_2020} found that, within the tropics,
\(ANPP_{stem}\) decreases at the highest maximum temperatures. Various
measures of primary productivity saturate at high levels of MAP, though
the saturation points identified vary from 1500mm
\citep{luyssaert_co_2007} up to 2,445 mm MAP
\citep{schuur_productivity_2003}. Interactions between MAT and MAP are
also possible; within the tropics, there is a positive interaction
between MAT and MAP in shaping \(ANPP\), such that high rainfall has a
negative effect on productivity in cooler climates, compared to a
positive effect in warmer climates \citep{taylor_temperature_2017}.
\textbf{Such complicated dynamics play out with belowground fluxes as
well, with typically dominant soil temperature mediated by moisture and
C supply \citep{hursh_sensitivity_2017, xu_contribution_2016}.
\url{https://github.com/forc-db/Global_Productivity/issues/111}} There
is also evidence that C fluxes also respond to climate variables such as
temperature and precipitation seasonality (\textbf{REFS}), cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{beer_terrestrial_2010, fyllas_solar_2017}, and potential
evapotranspiration \citep{kerkhoff_plant_2005}; however, these are not
typically assessed in global-scale analyses of annual forest C flux.

As metrics of annual climate, MAT and MAP fail to capture variation in
climate on an intra-annual scale, including temperature and
precipitation seasonality and growing season length. Some studies have
suggested that the apparently strong relationship between MAT and C
fluxes is actually a consequence of the correlation between MAT and
growing season length
\citep{kerkhoff_plant_2005, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al.~\citeyearpar{kerkhoff_plant_2005} and Michaletz et
al.~\citeyearpar{michaletz_convergence_2014} found no significant
relationship between growing season temperature and net primary
productivity (loosely defined to include a mix of \textbf{VARIABLES})
standardized to growing season length. While this suggests that the
influence of temperature may be limited to determining the length of the
frost-free growing season, analysis with clearly defined variables and
standardized forest ages would be necessary to test the the veracity and
generality of this hypothesis.

The recent development of the Global Forest Carbon database (ForC),
which synthesizes multiple variables and including records of stand
history
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018},
opens up the possibility for a standardized analysis of global scale
variation in multiple C fluxes and the principle climatic drivers of
these patterns. In order to approach this broad topic, we simplify the
major gaps in our knowledge to five broad questions and corresponding
hypotheses (Table 1). First, we ask how nine forest autotrophic carbon
fluxes in ForC vary with latitude. We then test how these fluxes relate
to MAT and MAP, and additionally how they respond to other, less
well-studied, climate variables. Finally, we consider the relationship
between C flux and seasonality, considering the role of seasonality in
explaining variation in carbon fluxes, and the influence of climate on C
flux standardized by growing season length.

@@

\caption{\label{tab:unnamed-chunk-4}Summary of research questions, corresponding hypotheses, and results. Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', and '-' indicates no significant relationship. Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.}

@@

\begin{tabular}[t]{llllllllllll}

@@

\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes} & \multicolumn{1}{c}{ } \\

@@

Questions and hypotheses (with related references) & Overall & $GPP$ & $NPP$ & $ANPP$ & $ANPP_{stem}$ & $ANPP_{foliage}$ & $BNPP$ & $BNPP_{fine.root}$ & $R_{auto}$ & $R_{root}$ & Support\\

@@

\multicolumn{5}{l}{\textbf{Q1. How do C fluxes vary with latitude?}}\\
\hspace{1em}H1.1. C fluxes decrease continuously with latitude.\textsuperscript{1}$^{,}$\textsuperscript{2}$^{,}$\textsuperscript{3}$^{,}$\textsuperscript{10} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\

@@

\multicolumn{5}{l}{\textbf{Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?}}\\
\hspace{1em}H2.1. C fluxes increase continuously with MAT.\textsuperscript{1}$^{,}$\textsuperscript{4}$^{,}$\textsuperscript{9} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H2.2. C fluxes increase with precipitation up to at least 2000 mm yr$^{-1}$.\textsuperscript{1}$^{,}$\textsuperscript{4} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H2.3. Temperature and precipitation interactively shape C fluxes.\textsuperscript{5} & (yes) & yes & yes & - & yes & - & yes & yes & yes & - & Fig. 3\\

@@

\multicolumn{5}{l}{\textbf{Q3. How are C fluxes related to other annual climate variables?}}\\
\hspace{1em}H3.1. C fluxes display a decelerating increase or unimodal relationship with PET. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H3.2. C fluxes display a decelerating increase or unimodal relationship with vapour pressure deficit. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
\hspace{1em}H3.3. C fluxes increase with solar radiation. & (yes) & yes & yes & yes & yes & yes & yes & yes & yes & - & Figs. S4, S5\\

@@

\multicolumn{5}{l}{\textbf{Q4. How does seasonality influence annual C fluxes?}}\\
\hspace{1em}H4.1. C fluxes decrease with temperature seasonality. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S6, S7\\
\hspace{1em}H4.2. C fluxes decrease with precipitation seasonality. & - & - & - & - & no & - & - & - & - & - & Figs. S6, S7\\
\hspace{1em}H4.3. C fluxes increase with growing season length.\textsuperscript{6}$^{,}$\textsuperscript{7}$^{,}$\textsuperscript{8} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S6, S7\\
\hspace{1em}H4.4. Growing season length is a better predictor of C fluxes than MAT.\textsuperscript{7}$^{,}$\textsuperscript{8} & (no) & no & no & no & - & no & no & no & no & no & Table S4\\

@@

\multicolumn{5}{l}{\textbf{Q5. When standardised by growing season length, how do annual C fluxes vary with climate?}}\\
\hspace{1em}H5.1. Growing season C fluxes increase with temperature.\textsuperscript{8} & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
\hspace{1em}H5.2. Growing season C fluxes increase with PET. & (yes) & yes & yes & - & yes & - & yes & yes & - & - & Figs. S8, S9\\
\hspace{1em}H5.3. Growing season C fluxes increase with precipitation. & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
\hspace{1em}H5.4. Growing season C fluxes increase with solar radiation. & (yes) & - & - & - & - & - & yes & yes & - & - & Figs. S8, S9\\

@@

\multicolumn{12}{l}{\textsuperscript{1} Luyssaert et al. (2007) \textsuperscript{2} Gillman et al. (2015) \textsuperscript{3} Simova and Storch (2017) \textsuperscript{4} Schuur (2003) \textsuperscript{5} Taylor et al. (2016) \textsuperscript{6} Malhi (2012) \textsuperscript{7} Michaletz et al. (2014) \textsuperscript{8} Chu et al. (2016) \textsuperscript{9} Piao et al. (2010) \textsuperscript{10} Huston \& Wolverton (2009)}\\

@@

\hypertarget{materials-and-methods}{%
\subsubsection{Materials and Methods}\label{materials-and-methods}}
